Automatically Proving Equivalence by Type-Safe Reflection

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Certainly the final version

Working with equivalences instead of equalities will bring one complication. These proofs of correctness that we will have to produce by hand will be slightly more complicated, as we won't be able to use Idris' ``rewrite" command, that enables to rewrite a subterm of a term by another one, provided that the two subterms are equal. More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use ``rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$. This is a classical problem of working within a setoid, and this problem can be slightly mitigated when the programming language offers a good support for rewriting terms in setoids. However, Idris isn't equipped with any automation for setoids, and everything will have to be done by hand. For this purpose, we will use the following lemma :

@@

The remaining part is to define the function \code{norm}. Each algebraic structure will have a function for reducing the reflected terms into their normal forms. The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities. Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.
As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equivalence between the initial real value (the index of the input term) and the final real value (the index of the produced term). Thus, the type of the reduction function for group is :

>>>>

Working with equivalences instead of equalities will bring one complication. These proofs of correctness that we will have to produce by hand will be slightly more complicated, as we won't be able to use Idris' ``rewrite" command, that enables to rewrite a subterm of a term by another one, provided that the two subterms are equal. More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use ``rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$. This is a classical problem of working within a setoid, which is slightly mitigated when the programming language offers a good support for rewriting terms in setoids. However, Idris isn't equipped with any support for setoids, and everything will have to be done by hand. For this reason, we define the lemma :

@@

The remaining part is to define the function \code{norm}. Each algebraic structure will have a function for reducing the reflected terms into their normal forms. The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities. Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.
As said before, the normalisation functions also produce the proof of equivalence between the initial real value (the index of the input term) and the final real value (the index of the produced term). Thus, the type of the reduction function for group is :

### 2017-03-31 paper-CICM-17/content/4-relatedWork.tex
Certainly the final version

Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there is usually no normal form. This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and heuristics for proof search starts. \\

>>>>

Finally, if we leave the ground of nice mathematical structures, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such systems, because there is usually no normal form. This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and proof search heuristics starts. \\

### 2017-03-31 paper-CICM-17/content/5-conclusions.tex
Certainly the final version

We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were using indexed types. This solution takes the form of a hierarchy of provers for equivalence in algebraic structures. These provers are generic in multiple ways: they work for many algebraic structures (semi-group, monoid, commutative monoid, group, commutative group, ring and semi-ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality). The implementation is modular and each prover reuses the prover of the structure from which it directly inherits. These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations as \code{adc\_lemma\_2}. Therefore, these provers enable the user to focus on the interesting proofs that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated.\\

>>>>

We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were using indexed types. This solution takes the form of a hierarchy of provers for equivalence in algebraic structures. These provers are generic in multiple ways: they work for many algebraic structures (semi-group, monoid, commutative monoid, group, commutative group, ring and semi-ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality). The implementation is modular and each prover reuses the prover of the structure from which it inherits. These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations, as \code{adc\_lemma\_2}. Thus, these provers enable the user to focus on the interesting proofs that requires specific knowledge and creativity, instead of wasting time for proving routine lemmas that can be automated.\\
